---
title: "Vignette for MLG: multi-layer graph clustering for scRNA-seq data"
author: "Shan Lu"
output: 
  rmarkdown::html_document:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{MLG Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup}
suppressPackageStartupMessages({
  library(RColorBrewer)
  library(ggplot2)
  library(cowplot)
  library(knitr)
  library(Seurat)
  library(liger)
  library(mlg)
})
data("kowalczyk_1", package = "mlg")
theme_set(theme_classic())
```

# Dataset description
- Three variables are stored in "kowalczyk_1": gene_expression_matrix, cell_info, cnmf_factor.
- "gene_expression_matrix" stores the logTPM of 1058 cells.
- "cell_info" stores the true cell type labels (LT-HSC, MPP, ST-HSC) and condition labels (old and young) of the cells.
- "cnmf_factor" stores the coordinates from cNMF decomposition. We include cNMF factors here, because cNMF package is implemented in python.

# Clustering analysis
## Perform PCA, tSNE and Louvain clustering with PCA factors.

```{r, fig.hold='hold', fig.width=8, fig.height=6}
# number of component to be kept for each dimension reduction method
num_component <- 15

# generate Seurat object
kowalczyk_seurat <- CreateSeuratObject(counts = gene_expression_matrix, project = "kowalczyk_seurat ", 
                              assay = "RNA",
                              meta.data = cell_info)

kowalczyk_seurat <- FindVariableFeatures(kowalczyk_seurat, selection.method = "vst", nfeatures = 3000, verbose = F)
kowalczyk_seurat <- ScaleData(kowalczyk_seurat, vars.to.regress = 'condition_label', verbose = F)
kowalczyk_seurat <- RunPCA(kowalczyk_seurat, npcs=num_component, verbose = F)
kowalczyk_seurat <- RunTSNE(kowalczyk_seurat, dims = 1:num_component, verbose = F)
kowalczyk_seurat <- FindNeighbors(kowalczyk_seurat, dims = 1:num_component, verbose = F)
kowalczyk_seurat <- FindClusters(kowalczyk_seurat, resolution = 0.2, algorithm =1, verbose = F)
tSNE_coord_PCA <- data.frame(kowalczyk_seurat@reductions$tsne@cell.embeddings)
# Visualization of the dataset on tSNE coordinates
plot_grid(ggplot(tSNE_coord_PCA,aes(tSNE_1, tSNE_2, color=cell_info$cell_type_label))+
            geom_point()+labs(color="cell type")+theme(legend.position = "bottom"),
          ggplot(tSNE_coord_PCA,aes(tSNE_1, tSNE_2, color=kowalczyk_seurat$seurat_clusters))+
            labs(color="cluster")+theme(legend.position = "bottom")+geom_point())
# The Adjusted Rand Index(ARI) of the Louvain-PCA clustering
pdfCluster::adj.rand.index(kowalczyk_seurat$seurat_clusters, cell_info$cell_type_label)

```


## Perform Seurat integration, tSNE and Louvain clustering with PCA components based on Seurat integrated data.

```{r, fig.hold='hold', fig.width=8, fig.height=6}
# Seurat integration
condition.list <- SplitObject(kowalczyk_seurat, split.by = "condition_label")
int.anchors <- FindIntegrationAnchors(object.list = condition.list, anchor.features=3000, verbose = F)
kowalczyk.integrated <- IntegrateData(anchorset = int.anchors, verbose = F)
DefaultAssay(kowalczyk.integrated) <- "integrated"
# Reorder the cells, so that the cell order in object kowalczyk.integrated is the same as in original dataset.
kowalczyk.integrated@assays$integrated@data<- kowalczyk.integrated@assays$integrated@data[, rownames(cell_info)]
kowalczyk.integrated@meta.data<-kowalczyk.integrated@meta.data[rownames(cell_info),]
kowalczyk.integrated <- ScaleData(kowalczyk.integrated, verbose = F)
kowalczyk.integrated <- RunPCA(kowalczyk.integrated, npcs = 30, verbose = F)
kowalczyk.integrated <- RunTSNE(kowalczyk.integrated, verbose = F)
kowalczyk.integrated <- FindNeighbors(kowalczyk.integrated,dims = 1:num_component, k.param = 20, verbose = F)
kowalczyk.integrated <- FindClusters(kowalczyk.integrated, resolution = 0.22, algorithm = 1, verbose = F)
tSNE_coord_Seurat <- data.frame(kowalczyk.integrated@reductions$tsne@cell.embeddings)
# Visualization of the dataset on tSNE coordinates
plot_grid(ggplot(tSNE_coord_Seurat,aes(tSNE_1, tSNE_2, color=cell_info$cell_type_label))+
            geom_point()+labs(color="cell type")+theme(legend.position = "bottom"),
          ggplot(tSNE_coord_Seurat,aes(tSNE_1, tSNE_2, color=kowalczyk.integrated$seurat_clusters))+
            labs(color="cluster")+theme(legend.position = "bottom")+geom_point())
# The Adjusted Rand Index(ARI) of the Seurat clustering
pdfCluster::adj.rand.index(kowalczyk.integrated$seurat_clusters, cell_info$cell_type_label)

```

## Perform Liger integration, tSNE and Louvain clustering with Liger factors.

```{r, fig.hold='hold', fig.width=8, fig.height=6, message=F, warning=F}

ligerex = createLiger(list(young=condition.list[[1]]@assays$RNA@counts, 
                           old=condition.list[[2]]@assays$RNA@counts)) 
ligerex = liger::normalize(ligerex)
ligerex = selectGenes(ligerex, var.thresh = .7281) # keep 3000 genes
length(ligerex@var.genes)
ligerex = scaleNotCenter(ligerex)

ligerex = optimizeALS(ligerex, k = num_component) 
ligerex = quantileAlignSNF(ligerex, resolution=0.4)
ligerex<- runTSNE(ligerex)

tSNE_coord_Liger <- data.frame(ligerex@tsne.coords)
colnames(tSNE_coord_Liger) <- c("tSNE_1", "tSNE_2")
# Visualization of the dataset on tSNE coordinates
plot_grid(ggplot(tSNE_coord_Liger,aes(tSNE_1, tSNE_2, color=cell_info$cell_type_label))+
            geom_point()+labs(color="cell type")+theme(legend.position = "bottom"),
          ggplot(tSNE_coord_Liger,aes(tSNE_1, tSNE_2, color=ligerex@clusters))+
            labs(color="cluster")+theme(legend.position = "bottom")+geom_point())
# The Adjusted Rand Index(ARI) of the Liger clustering
pdfCluster::adj.rand.index(ligerex@clusters, cell_info$cell_type_label)

```

## Perform cNMF decomposition, tSNE and Louvain clustering with cNMF factors.

```{r, fig.hold='hold', fig.width=8, fig.height=6,  message=F, warning=F}

cnmf_graph <- FindNeighbors(cnmf_factor, verbose = F)
cnmf_cluster <- FindClusters(cnmf_graph$snn, resolution = 0.15, algorithm =1, verbose = F)[[1]]
tSNE_cNMF<-RunTSNE(cnmf_factor, verbose=F)
tSNE_coord_cNMF =data.frame(tSNE_cNMF@cell.embeddings)
# Visualization of the dataset on tSNE coordinates
plot_grid(ggplot(tSNE_coord_cNMF,aes(tSNE_1, tSNE_2, color=cell_info$cell_type_label))+
            geom_point()+labs(color="cell type")+theme(legend.position = "bottom"),
          ggplot(tSNE_coord_cNMF,aes(tSNE_1, tSNE_2, color=cnmf_cluster))+
            labs(color="cluster")+theme(legend.position = "bottom")+geom_point())
# The Adjusted Rand Index(ARI) of the Louvain-cNMF clustering
pdfCluster::adj.rand.index(cnmf_cluster, cell_info$cell_type_label)

```


## MLG clustering

```{r, fig.hold='hold', fig.width=10, fig.height=10}

# reordering and scaling of raw liger factors
liger_factor=rbind(ligerex@H$young,ligerex@H$old)
# reorder the cells, so that the cell order are the same as in original data
liger_factor=liger_factor[rownames(cell_info),]
# scale the penalized NMF score for each cell, so that they sums to 1
liger_factor=t(apply(liger_factor, 1, function(i){i/sum(i)}))

factor_list<- list(PCA=kowalczyk_seurat@reductions$pca@cell.embeddings[,1:num_component],
                   cNMF=cnmf_factor,
                   Seurat=kowalczyk.integrated@reductions$pca@cell.embeddings[,1:num_component],
                   Liger=liger_factor)
# perform MLG clustering 
MLG = mlg_cluster(factor.list = factor_list, cluster.resolution=.5)

# The Adjusted Rand Index(ARI) of the MLG clustering
pdfCluster::adj.rand.index(MLG, cell_info$cell_type_label)

# visualization of the overlapping (which characterizes dependences) between different SNN layers.
p = prop.overlap.edges(factor_list)
p

# compute graph signal to noice ratio of the SNN graph constructed from PCA, cNMF, Seurat, Liger factors and MLG.
graph_signal_noise_ratio(factor_list, knn.param=20, prune.param=1/5, cell_label=cell_info$cell_type_label)

# compare clustering ARI of the 5 methods.
clustering_acc = c(pdfCluster::adj.rand.index(kowalczyk_seurat$seurat_clusters, cell_info$cell_type_label),
  pdfCluster::adj.rand.index(cnmf_cluster, cell_info$cell_type_label),
  pdfCluster::adj.rand.index(kowalczyk.integrated$seurat_clusters, cell_info$cell_type_label),
  pdfCluster::adj.rand.index(ligerex@clusters, cell_info$cell_type_label),
  pdfCluster::adj.rand.index(MLG, cell_info$cell_type_label))
names(clustering_acc) = c("PCA", "cNMF", "Seurat", "Liger", "MLG")
print(clustering_acc)
```



# Session

```{r}
sessionInfo()
```
